---
title: "Reproducible Research Assignment1"
author: "Vivek"
date: "August 16, 2015"
output: html_document
---

#### Loading and preprocessing the data

```{r,echo = TRUE}
setwd("C:/Users/Vivek/Desktop/Anaytics/Coursera/cleaning_data")
act<-read.csv("activity.csv",stringsAsFactors =FALSE)
act$date <- as.Date(act$date)
```


#### What is mean total number of steps taken per day?

Histogram of Total Steps

```{r, echo=TRUE}
steps<-aggregate(act$steps ~act$date, FUN = sum )
colnames(steps)<-c("day","steps")
hist(steps$steps,xlab="Steps",main= "Histogram of Total Steps")

```

Mean and Median Number of steps taken

```{r, echo=TRUE}

mean(steps$steps)
median(steps$steps)
```

#### What is the average daily activity pattern?

```{r, echo=TRUE}

time_series<-aggregate(act$steps ~act$interval, FUN=mean)
colnames(time_series)<-c("Interval","Avg_Steps")
plot(time_series$Interval,time_series$Avg_Steps,type="l",xlab="Interval",ylab="Average Steps")
```

The 5-minute interval, which on average across all the days in the dataset, contains the maximum number of steps is 

```{r, echo=TRUE}

head(time_series[with(time_series, order(-Avg_Steps)), ],1)
```

#### Imputing missing 

The total number of rows containing missing values are 2304.

```{r, echo=TRUE}

summary(act)
```

#### Making a new Dataset with imputed missing values

For this the missing value is replaced by the mean for the 5-minute interval, for that particular observation

```{r, echo=TRUE}

mean_steps<-aggregate(act$steps ~act$interval, FUN=mean)
colnames(mean_steps)<-c("interval","steps")
steps2<-act
steps2_null<-subset(steps2,is.na(steps2$steps))
steps2_not_null<-subset(steps2,!is.na(steps2$steps)) 
step2_imputed<-merge(steps2_null,mean_steps,by="interval",all.x=TRUE)
step2_imputed$steps.x<-NULL
colnames(step2_imputed)<-c("interval","date","steps")
step2_imputed<-step2_imputed[,c(3,2,1)]
steps2_final<-rbind(steps2_not_null,step2_imputed)
```

 Histogram of the total number of steps taken each day

```{r, echo=TRUE}

steps_total<-aggregate(steps2_final$steps ~steps2_final$date, FUN = sum )
colnames(steps_total)<-c("day","steps")
hist(steps_total$steps,xlab="Steps",main= "Histogram of Total Steps")
```

Mean and Median Number of steps taken in new dataset

```{r, echo=TRUE}
mean(steps_total$steps)
median(steps_total$steps)
```
There is no major difference between the mean and median of original dataset with imputed dataset.

#### Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}
steps2_final$day <-weekdays(steps2_final$date)

steps2_final$day<-ifelse(steps2_final$day =="Saturday"|steps2_final$day =="Sunday" ,"weekend","weekday")
avgSteps <- aggregate(steps2_final$steps, 
                      list(interval = as.numeric(as.character(steps2_final$interval)), 
                           day = steps2_final$day),
                      FUN = "mean")
names(avgSteps)[3] <- "mean_steps"

library(lattice)
xyplot(avgSteps$mean_steps ~ avgSteps$interval | avgSteps$day, 
       layout = c(1, 2), type = "l", 
       xlab = "Interval", ylab = "Number of steps")


```